---
title: "Paper"
author: "Jinxuan Lu"
date: "11/28/2021"
output:
  html_document: 
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

#import data and loading packages

```{r}
Sys.setenv(LANG="en")
getwd()
library(tidyverse)
library(haven)
library(dplyr)
library(psych)
library(knitr)

#import data
#ESS8 <- read_sav("Desktop/MyProject/ESS8e02_2.sav")
ESS8 <- read_sav("/Users/dinolu/Desktop/Finished courses/SIMM61-Quantitative Data Analysis in R/Individual paper相关/ESS8e02_2.sav")

#look at the data
View(ESS8)
unique(ESS8$cntry)

# Filter data by age range
ESS8_sample <- ESS8 %>% 
  filter(agea<=80 & agea >= 15)
View(ESS8_sample)
```

#mutate and select variables that fit for my research

```{r}
#Mutate and select variables
ESS8_Selected <- ESS8_sample %>%
  mutate(Age = as.numeric(agea),
         Gender = as_factor(gndr),
         Manage_income = as_factor(hincfel),
         Live_with_partner = as_factor(icpart2),
         Live_with_children = as_factor(chldhm),
         Work_agency_type = as_factor(tporgwk),
         Domicile = as_factor(domicil),
         Climate_change_concern = as.numeric(wrclmch),
         Self_efficacy = as.numeric(ownrdcc),
         Collective_efficacy = as.numeric(lkredcc),
         Generalized_trust = as.numeric(lklmten),
         Institutional_trust = as.numeric(gvsrdcc),
         Personal_norm = as.numeric(ccrdprs)) %>%
  select(Age,
         Gender,
         Manage_income,
         Live_with_partner,
         Live_with_children,
         Work_agency_type,
         Domicile,
         Climate_change_concern,
         Self_efficacy,
         Collective_efficacy,
         Generalized_trust,
         Institutional_trust,
         Personal_norm) %>% filter(Climate_change_concern<6, Self_efficacy<50, Collective_efficacy<50, Generalized_trust<50, Institutional_trust<50, Personal_norm<50)

view(ESS8_Selected)


#remove missing values
ESS8_Selected <- drop_na(ESS8_Selected)
view(ESS8_Selected)
```

#deal with categorical control variables in advance

```{r}
# Reduce the number of categories in a factor
ESS8_Selected <- ESS8_Selected %>% 
  mutate(Work_agency_type = fct_collapse(Work_agency_type,
                                  gov = c("Central or local government",
                                          "Other public sector (such as education and health)"),
                                  ente = c("A state owned enterprise",
                                           "A private firm"),
                                  other = c("Self employed",
                                            "Other")),
         Domicile = fct_collapse(Domicile, 
                                BC = c("A big city",
                                       "Suburbs or outskirts of big city"), 
                                Other = c("Town or small city", 
                                          "Country village", 
                                          "Farm or home in countryside")),
         Manage_income = fct_collapse(Manage_income, 
                                      Manage = c("Living comfortably on present income",
                                                 "Coping on present income"),
                                      Not = c("Difficult on present income",
                                              "Very difficult on present income"))) 
view(ESS8_Selected)
```

#set different dummy variables

```{r}
# Relevel to set reference category
ESS8_Selected <- ESS8_Selected %>% 
  mutate(Gender = fct_relevel(Gender, "Female"),
         Live_with_partner = fct_relevel(Live_with_partner, "Does not"),
         Live_with_children = fct_relevel(Live_with_children, "Does not"),
         Work_agency_type = fct_relevel(Work_agency_type, "other"),
         Manage_income = fct_relevel(Manage_income, "Not"),
         Domicile = fct_relevel(Domicile, "Other"))
view(ESS8_Selected)
```

#index is not available in my study, so no index included. 
#test bivariate model

```{r}
# Bivariate model 
mod1 <- lm(Personal_norm ~ Collective_efficacy,
           data = ESS8_Selected)

summary(mod1)
```

#add control variables

```{r}
# Model with dummies
mod2 <- lm(Personal_norm ~ Collective_efficacy +
             Age +
             Gender +
             Manage_income +
             Live_with_partner +
             Live_with_children +
             Work_agency_type +
             Domicile +
             Climate_change_concern +
             Self_efficacy, 
           data = ESS8_Selected)

summary(mod2)
```

#adding interaction terms to execute moderating analysis

```{r}
# Model with interaction effects
# moderator 1: Generalized trust
mod3 <- lm(Personal_norm ~ Collective_efficacy +
             Age +
             Gender +
             Manage_income +
             Live_with_partner +
             Live_with_children +
             Work_agency_type +
             Domicile +
             Climate_change_concern +
             Self_efficacy +
             Generalized_trust +
             Collective_efficacy * Generalized_trust,
           data = ESS8_Selected)

summary(mod3)
```

#adding interaction terms to execute moderating analysis

```{r}
# Model with interaction effects
# moderator 2: Institutional trust
mod4 <- lm(Personal_norm ~ Collective_efficacy +
             Age +
             Gender +
             Manage_income +
             Live_with_partner +
             Live_with_children +
             Work_agency_type +
             Domicile +
             Climate_change_concern +
             Self_efficacy +
             Institutional_trust +
             Collective_efficacy * Institutional_trust,
           data = ESS8_Selected)

summary(mod4) 
#now it seems only moderator 1 (mod3) plays a role
```

#run plots to check my models

```{r}
# Detecting influential outliers
plot(mod3)
```

#deal with outliers

```{r}
# display and exclude the outlier cases through Cook's D
# I searched and found that a general rule of thumb is to investigate any point that is more than 3*mean of all the distances
CooksD <- cooks.distance(mod3)
influential <- CooksD[(CooksD > (3 * mean(CooksD, na.rm = TRUE)))]

names_of_influential <- names(influential)
outliers <- ESS8_Selected[names_of_influential,]
ESS8_Selected_without_outliers <- ESS8_Selected %>% anti_join(outliers)
view(ESS8_Selected_without_outliers)
#A new final sample called ESS8_Selected_without_outliers including 31670 cases.
```

#official analysis with final sample data \#means, standard deviation, and correlation among numeric variables

```{r}
library(gtsummary)
ESS8_Selected_without_outliers %>% tbl_summary()
library(apaTables)
apa.cor.table(ESS8_Selected_without_outliers,
              show.conf.interval = TRUE,
              show.sig.stars = TRUE,
              landscape = TRUE)
```

#visualize the focal variables

```{r}
# Plot focal X
ggplot(ESS8_Selected_without_outliers, aes(Collective_efficacy)) + 
  geom_histogram() +
  theme_classic() + 
  ggtitle("Focal Predictor Variable")

# Summary stats for focal X 
summary(ESS8_Selected_without_outliers$Collective_efficacy)
```

```{r}
# Plot focal Y
ggplot(ESS8_Selected_without_outliers, aes(Personal_norm)) + 
  geom_histogram() +
  theme_classic() + 
  ggtitle("Focal Outcome Variable")

# Summary stats for focal X 
summary(ESS8_Selected_without_outliers$Personal_norm)
```

#look at the whole picture

```{r}
library(skimr)
skim(ESS8_Selected_without_outliers) 
```

#rerun the models

```{r}
# model 1, focal relationship
mod1 <- lm(Personal_norm ~ Collective_efficacy,
           data = ESS8_Selected_without_outliers)
summary(mod1)

```

```{r}
# model 2, adding control variables into focal relationship
mod2 <- lm(Personal_norm ~ Collective_efficacy +
             Age +
             Gender +
             Manage_income +
             Live_with_partner +
             Live_with_children +
             Work_agency_type +
             Domicile +
             Climate_change_concern +
             Self_efficacy, 
           data = ESS8_Selected_without_outliers)
summary(mod2)
```

```{r}
# model 3 with interaction effects
# moderator 1: Generalized Trust
mod3 <- lm(Personal_norm ~ Collective_efficacy +
             Age +
             Gender +
             Manage_income +
             Live_with_partner +
             Live_with_children +
             Work_agency_type +
             Domicile +
             Climate_change_concern +
             Self_efficacy +
             Generalized_trust +
             Collective_efficacy * Generalized_trust,
           data = ESS8_Selected_without_outliers)
summary(mod3)
```

```{r}
# model 4 with interaction effects
# moderator 2: Institutional Trust
mod4 <- lm(Personal_norm ~ Collective_efficacy +
             Age +
             Gender +
             Manage_income +
             Live_with_partner +
             Live_with_children +
             Work_agency_type +
             Domicile +
             Climate_change_concern +
             Self_efficacy +
             Institutional_trust +
             Collective_efficacy * Institutional_trust,
           data = ESS8_Selected_without_outliers)
summary(mod4) 
```

#exclude the independent variable and only include control variables to get the delta R2

```{r}
mod5 <- lm(Personal_norm ~ Age +
             Gender +
             Manage_income +
             Live_with_partner +
             Live_with_children +
             Work_agency_type +
             Domicile +
             Climate_change_concern +
             Self_efficacy, 
           data = ESS8_Selected_without_outliers)
summary(mod5)
```

#visualize the final model

```{r}
library(visreg)
#visualize focal relationship
visreg(mod2, "Collective_efficacy", 
       ylab = "Personal Norm toward Climate Change", 
       xlab = "Collective Efficacy Belief", 
       gg = TRUE, # creating a ggplot
       band = TRUE) + 
  theme_classic() + 
  ggtitle("Final focal model")

```

```{r}
# visualize moderation relationship
visreg(mod3, "Collective_efficacy", by = "Generalized_trust", 
                      overlay = TRUE,
                      ylab = "Personal Norm toward Climate Change", 
                      xlab = "Collective Efficacy Belief",
                      legend = FALSE,
                      gg = TRUE, 
       band = FALSE) + 
  theme_classic() + 
  ggtitle("Final model 1: The interaction effect of collective efficacy and generalized trust")

library(interactions)
interact_plot(mod3, pred = Collective_efficacy, modx = Generalized_trust)
#it seems the latter plot is better for continuous moderators
```

```{r}
# visualize moderation relationship
visreg(mod4, "Collective_efficacy", by = "Institutional_trust", 
                      overlay = TRUE,
                      ylab = "Personal Norm toward Climate Change", 
                      xlab = "Collective Efficacy Belief",
                      legend = FALSE,
                      gg = TRUE, 
       band = FALSE) + 
  theme_classic() + 
  ggtitle("Final model 2: The interaction effect of collective efficacy and institutional trust")

interact_plot(mod4, pred = Collective_efficacy, modx = Institutional_trust)

```

#model summary

```{r}
library(modelsummary)
msummary(list(mod1, mod2, mod3, mod4, mod5), # List the models to include
         stars = TRUE, statistic = NULL)

```

